---
title: "R Notebook"
output: 
--- 

```{r}
if (!require("ggplot2")) { install.packages("ggplot2"); require("ggplot2") }
if (!require("evd")) { install.packages("evd"); require("evd") }
if (!require("evir")) { install.packages("evir"); require("evir") }
if (!require("nortest")) { install.packages("nortest"); require("nortest") }
if (!require("insuranceData")) { install.packages("insuranceData"); require("insuranceData") }
if (!require("extRemes")) { install.packages("extRemes"); require("extRemes") }
if (!require("VaRES")) { install.packages("VaRES"); require("VaRES") }
if (!require("vars")) { install.packages("vars"); require("vars") }
if (!require("PerformanceAnalytics")) { install.packages("PerformanceAnalytics"); require("PerformanceAnalytics") }
```


```{r}
library(ggplot2)
library(evd)
library(evir)
library(nortest)
library(insuranceData)
library(extRemes)
library(VaRES)
library(vars)
library(PerformanceAnalytics)
```

The AutoClaims dataset contains data on claims experience from a large midwestern (US) P&C insurer for private motor insurance. The dependent variable is the amount paid on a closed claim, in $. Claims that were not closed by year end are handled separately.

```{r}
data(AutoClaims)
dataset <- AutoClaims$PAID
summary(dataset)
hist(dataset)
```
Block Maxima
```{r}
blocks <- blockmaxxer.data.frame(data.frame(dataset), blen=200, span=33)
blocks
hist(blocks)
```

Anderson-Darling Test of robustness of the data

```{r}
ad.test(blocks)
```
gev: maximum likelihood estimation of the 3-parameter generalized extreme value (GEV) distribution. 
fgev: maximum-likelihood fitting for the generalized extreme value distribution, including linear modelling of the location parameter, and allowing any of the parameters to be held fixed if desired.
```{r}
GEV <- gev(blocks)
fit <- fgev(GEV$data, std.err = FALSE)
plot.gev(GEV)
fit
epsilon = fit$estimate[3]
sigma = fit$estimate[2]
mu = fit$estimate[1]
print(mu)
```

Since shape parameter > 0, the data follows a Fréchet distribution.
Plot
```{r}
par(mfrow=c(1,2))
plot(density(blocks), main="Density Function of AutoClaims")
x <- seq(-10000, 70000, by=1)
Frechet_density <- dgev(x, mu, sigma, epsilon)
#Since NaNs are produced, we must used na.omit to remove these values
plot(na.omit(Frechet_density), main="Density Function of Fitted Frechet")
```

```{r}
FrechetDistribution <- data.frame(x, Frechet_density)
names(FrechetDistribution) <- c("Time", "Observations")
frechet_var_ts <- ts(na.omit(FrechetDistribution))
var.2c <- VAR(frechet_var_ts, p = 2, type = "const")

# Diagnostic Testing
# Arch Test
arch.test(var.2c)
# Normality Test
normality.test(var.2c)
# Predict the values for the next 8 days
var.2c.prd <- predict(var.2c, n.ahead = 8, ci = 0.9)
plot(var.2c.prd)
# The conditional drawdown is the the mean of the worst 0.99% drawdowns
CDD(frechet_var_ts, weights = NULL, geometric = TRUE, invert = TRUE, p = 0.99)
# Calculates Expected Shortfall (also known as) Conditional Value at Risk
# Using modified Cornish Fisher calc for non-normal distribution
ES(frechet_var_ts, p=.99, method="modified")
```


Calculating the Expected Shortfall
```{r}
p = 0.99
vargev(p, mu, sigma, epsilon)
esgev(p, mu, sigma, epsilon)
```
```{r}
quantile(blocks,0.99)
mean(blocks[blocks>=quantile(dataset,0.99)])
mean(blocks)
```
