---
title: "R Notebook"
output: html_notebook
--- 

```{r}
if (!require("ggplot2")) { install.packages("ggplot2"); require("ggplot2") }
if (!require("evd")) { install.packages("evd"); require("evd") }
if (!require("evir")) { install.packages("evir"); require("evir") }
if (!require("nortest")) { install.packages("nortest"); require("nortest") }
if (!require("insuranceData")) { install.packages("insuranceData"); require("insuranceData") }
if (!require("extRemes")) { install.packages("extRemes"); require("extRemes") }
if (!require("VaRES")) { install.packages("VaRES"); require("VaRES") }
```


```{r}
library(ggplot2)
library(evd)
library(evir)
library(nortest)
library(insuranceData)
library(extRemes)
library(VaRES)
```

The AutoClaims dataset contains data on claims experience from a large midwestern (US) P&C insurer for private motor insurance. The dependent variable is the amount paid on a closed claim, in $. Claims that were not closed by year end are handled separately.

```{r}
data(AutoClaims)
dataset <- AutoClaims$PAID
summary(dataset)
hist(dataset)
```
Block Maxima
```{r}
blocks <- blockmaxxer.data.frame(data.frame(dataset), blen=100, span=67)
blocks
hist(blocks)
```

Anderson-Darling Test of robustness of the data

```{r}
ad.test(blocks)
```
gev: maximum likelihood estimation of the 3-parameter generalized extreme value (GEV) distribution. 
fgev: maximum-likelihood fitting for the generalized extreme value distribution, including linear modelling of the location parameter, and allowing any of the parameters to be held fixed if desired.
```{r}
GEV <- gev(blocks)
fit <- fgev(GEV$data, std.err = FALSE)
plot.gev(GEV)
fit
epsilon = fit$estimate[3]
sigma = fit$estimate[2]
mu = fit$estimate[1]
```

Since shape parameter > 0, the data follows a Fréchet distribution.
Plot
```{r}
par(mfrow=c(1,2))
plot(density(blocks), main="Density Function of AutoClaims")
x <- seq(-10000, 70000, by=1)
Frechet_density <- dgev(x, mu, sigma, epsilon)
#Since NaNs are produced, we must used na.omit to remove these values
Frechet_density <- na.omit(Frechet_density)
plot(Frechet_density, main="Density Function of Fitted Frechet")
```

Calculating the Expected Shortfall
```{r}
p = 0.99
#vargev(p, mu, sigma, epsilon)
esgev(p, mu, sigma, epsilon)
```
