---
title: "R Notebook"
output: 
--- 

```{r}
if (!require("ggplot2")) { install.packages("ggplot2"); require("ggplot2") }
if (!require("evd")) { install.packages("evd"); require("evd") }
if (!require("evir")) { install.packages("evir"); require("evir") }
if (!require("nortest")) { install.packages("nortest"); require("nortest") }
if (!require("insuranceData")) { install.packages("insuranceData"); require("insuranceData") }
if (!require("extRemes")) { install.packages("extRemes"); require("extRemes") }
if (!require("VaRES")) { install.packages("VaRES"); require("VaRES") }
if (!require("vars")) { install.packages("vars"); require("vars") }
if (!require("PerformanceAnalytics")) { install.packages("PerformanceAnalytics"); require("PerformanceAnalytics") }
if (!require("fExtremes")) { install.packages("fExtremes"); require("fExtremes") }
if (!require("RobExtremes")) { install.packages("RobExtremes"); require("RobExtremes") }
```


```{r}
library(ggplot2)
library(evd)
library(evir)
library(nortest)
library(insuranceData)
library(extRemes)
library(VaRES)
library(vars)
library(PerformanceAnalytics)
library(fExtremes)
library(RobExtremes)
```

The AutoClaims dataset contains data on claims experience from a large midwestern (US) P&C insurer for private motor insurance. The dependent variable is the amount paid on a closed claim, in $. Claims that were not closed by year end are handled separately.

```{r}
data(AutoClaims)
dataset <- AutoClaims$PAID
summary(dataset)
hist(dataset)
```
Block Maxima
```{r}
blocks <- blockmaxxer.data.frame(data.frame(dataset), blen=50, span=135)
blocks
hist(blocks)
```

Anderson-Darling Test of robustness of the data

```{r}
ad.test(blocks)
```
gevFit estimate the parameters either by the probability weighted moment method("pwm") or by maximum log likelihood estimation ("mle"). We have used PWM method here.
```{r}
fittedGEV <- gevFit(dataset, block = 200, type = c("pwm"))
epsilon = fittedGEV@fit$par.ests[1]
sigma = fittedGEV@fit$par.ests[3]
mu = fittedGEV@fit$par.ests[2]
epsilon
summary(fittedGEV)
```

Since shape parameter > 0, the data follows a Fréchet distribution.
Plot
```{r}
par(mfrow=c(1,2))
plot(density(blocks), main="Density Function of AutoClaims")
x <- seq(-10000, 70000, by=1)
Frechet_density <- dgev(x, mu, sigma, epsilon)
#Since NaNs are produced, we must used na.omit to remove these values
plot(na.omit(Frechet_density), main="Density Function of Fitted Frechet")
```

Calculating the Expected Shortfall
```{r}
p = 0.99
esgev(p, mu, sigma, epsilon)
```
```{r}
mean(blocks[blocks>=quantile(dataset,0.99)])
```
